---
title: "AEIS Data Example -- Generating a Synthetic RCT"
author: "Charlotte Mann"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(optmatch)
library(dplyr)
library(tidyr)
library(kableExtra)
```

## Load Data

```{r}
load("../input/MS_data_public.Rdata")
var.names.clean <- read.csv("../data/var_names.csv")
```


## Clean some variable names for interpretability

```{r}
covs_ms <- covs_ms %>%
  rename(all_stud_n = CPETALLC_67, grade8_n = CPETG08C_67, stud_teach_rat = CPSTKIDR_67,
         all_exp = CPFPAALLT_67, inst_exp = CPFEAINSP_67, lead_exp = CPFEAADSP_67, supp_exp = CPFEASUPP_67,
         ed_aide = CPSETOFC_67, teach_salary = CPSTTOSA_67, 
         teach_expr = CPSTEXPA_67, perc_teach_female = CPSTFEFP_67,
         perc_teach_white = CPSTWHFP_67, perc_teach_black = CPSTBLFP_67, perc_teach_hisp = CPSTHIFP_67,
         perc_stud_black = CPETBLAP_67, perc_stud_hisp = CPETHISP_67, 
         perc_stud_api = CPETPACP_67, perc_stud_white = CPETWHIP_67,
         perc_stud_alp = CPETSPEP_67, perc_stud_bil = CPETBILP_67, perc_stud_tag = CPETGIFP_67) %>%
  select(CAMPUS, GRDSPAN, starts_with("pre"),all_of(var.names.clean$var_clean), everything())
```

```{r, echo = F}
var.names.clean %>%
  kable(col.names = c("Original Variable", "Variable Description", "Clean Name"),
                     align = "lll", format = "html") %>% 
  row_spec(row = 0, align = "c", bold=T) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

## Form Fake/Example Paired RCT

First, match pairs of schools on baseline characteristics. Note, that this is not how one would design a paired trial because I used a bipartite pair matching algorithm from `optmatch` (which is just the package that I know how to use).

This results in 146 pairs of schools based on their previous 8th grade Math TAKS performance, size of the school, student and teacher demographics, and total school funding.

```{r}
set.seed(25)

# create a random grouping to be able to do bipartite matching
covs_ms$Z <- sample(0:1, nrow(covs_ms), replace = T)

# calculate distances between schools
dist_mat <- match_on(Z ~ premA + premB + premF + all_stud_n + stud_teach_rat + all_exp + 
                       perc_teach_white + perc_stud_hisp + 
                       perc_stud_alp + perc_stud_bil, data = covs_ms, caliper = 1)
summary(dist_mat)

# exclude distances outside of the caliper
initialmatch <- fullmatch(dist_mat, data = covs_ms)
excl <- which(is.na(initialmatch))

# and recalculate distances
dist.update <- match_on(Z ~ premA + premB + premF + all_stud_n + stud_teach_rat + all_exp + 
                       perc_teach_white + perc_stud_hisp + 
                       perc_stud_alp + perc_stud_bil, dat = covs_ms[-excl,]) 

# final optimal pair matching
fakematch <- pairmatch(dist.update,  data = covs_ms)
summary(fakematch)
```

Select 25 random pairs from those matched to make up our paired trial.

```{r}
set.seed(125)

match.ids <- unique(fakematch[which(!is.na(fakematch))])
rct.ids <- sample(match.ids, 25)
```

Randomly assign one school within each pair to receive treatment and save RCT information.

```{r}
rct <- covs_ms %>%
  select(CAMPUS) %>%
  mutate(match = fakematch) %>%
  filter(match %in% rct.ids) %>%
  mutate(match = factor(match,levels = rct.ids, labels = LETTERS[1:25])) %>%
  group_by(match) %>% 
  mutate(k = row_number()) %>%
  pivot_wider(values_from = CAMPUS, names_from = k, names_prefix = "CAMPUS") %>%
  mutate(Tr1 = sample(0:1,1),
         Tr2 = 1-Tr1) %>%
  pivot_longer(!match,
               names_to = c(".value","k"),
               names_pattern = "(.*)(1|2)") %>%
  select(CAMPUS, match, k, Tr) %>%
  arrange(match, k)

save(rct, file = "../data/rct_schools.Rdata")
```


Example Paired RCT schools:

```{r, echo = F}
rct %>%
  select(-k) %>%
  kable(format = "html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover","condensed"))

```

## Save Cleaned Data

```{r}
covs_rct = covs_ms %>%
  filter(CAMPUS %in% rct$CAMPUS)
covs_aux = covs_ms %>%
  filter(!(CAMPUS %in% rct$CAMPUS))

out_rct = rct %>%
  left_join(out_ms, by = "CAMPUS")
out_aux = out_ms %>%
  filter(!(CAMPUS %in% rct$CAMPUS))

save(out_rct, covs_rct, out_aux, covs_aux, file = "../data/rct_aux_dat.Rdata")
```