---
title: "Treatment Effect Estimation with dRCT"
author: "Charlotte Mann"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(kableExtra)
library(devtools)
```

## Load Data

Load data cleaned in `00-data-setup.Rmd`

```{r}
load("../data/rct_aux_dat.Rdata")
load("../data/rct_schools.Rdata")
```

## Install `dRCT' package

```{r}
if(!("dRCT" %in% installed.packages())) install_github("manncz/dRCT/dRCT")
library(dRCT)
```

## Estimate the ATE with Bernoulli Randomized Trails (LOOP)

Let's start only adjusting by the 2007 overall math TAKS passing rate `premA` using OLS for LOO potential outcome imputation.

```{r}
loop(Y = rct_dat$taks08, Tr = rct_dat$TrBern, Z = rct_dat$premA, pred = loop_ols)
```

Now, we can adjust with all covariates using Random Forest.

First, we need to carefully only chose pre-treatment covariates to adjust with.
```{r}
covMat = rct_dat %>%
  select(-CAMPUS, -GRDSPAN, -match, -k, -Tr, -TrBern, -taks08, -clust_size) %>%
  select(!starts_with("out"))
```


Then, we can input our matrix of pre-treatment covariates into the loop function:
```{r}
loop(Y = rct_dat$taks08, Tr = rct_dat$TrBern, Z = covMat, pred = loop_rf)
```

Compare these to the difference in means estimator.

```{r}
summary(lm(taks08 ~ TrBern, data = rct_dat))
```

## Estimate the ATE with a Paired Trial (P-LOOP)

```{r}
p_loop(Y = rct_dat$taks08, Tr = rct_dat$Tr, Z = rct_dat$premA, P = rct_dat$match, pred = p_ols_interp)
```



Difference in Means: 

```{r}
p_loop(Y = rct_dat$taks08, Tr = rct_dat$Tr, P = rct_dat$match)
```

## Estimate the ATE with a Paired Cluster Randomized Trial (IDPD)


```{r}
p_loop(Y = rct_dat$taks08, Tr = rct_dat$Tr, Z = rct_dat$premA, P = rct_dat$match, pred = p_ols_interp,
       n = rct_dat$clust_size)
```

```{r}
p_loop(Y = rct_dat$taks08, Tr = rct_dat$Tr, P = rct_dat$match, pred = p_ols_interp, n = rct_dat$clust_size)
```
