---
title: "Treatment Effect Estimation Integrating Auxiliary Data with `dRCT` Package"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
      out_dir <- "compiled_tutorials";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install packages if not already installed
needed.packages <- c("dplyr", "tidyr", "devtools","randomForest")
new.packages <- needed.packages[!(needed.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

```{r, message = FALSE}
library(dplyr)
library(tidyr)
library(randomForest)
library(devtools)
if(!("dRCT" %in% installed.packages())) install_github("manncz/dRCT")
library(dRCT)
```


## Overview

In this Markdown, we will walk through the steps outlined in [Gagnon-Bartsch et al. (2023)](https://www.degruyter.com/document/doi/10.1515/jci-2022-0011/html) for the RE-LOOP estimator, combining experimental and observational data for estimating the ATE with improved efficiency. 

We will continue to work with the synthetic RCT and AEIS data processed in `00-data-setup.Rmd`.

```{r}
load("../data/rct_aux_dat.Rdata")
```

## Step 1: Fit Auxiliary Model

We will fit the simplest auxiliary model possible -- using all available covariates in the auxiliary data with random forests. As a reminder, this data includes all of the Texas middle schools in the AEIS data that are not in the synthetic RCT. We use the `randomForest` package with the default parameters.

In practice, you can go to town developing an auxiliary model before looking at the RCT data. The only goal of the model is to predict the outcomes in the RCT well.

```{r, eval = F}
aux_mod_dat <- aux_dat %>%
  select(-GRDSPAN) %>%
  select(!starts_with("out"))

# we do not recommend re-running the auxiliary model because it takes a long time to run
#aux.rf <- randomForest(taks08 ~ . -CAMPUS, data = aux_mod_dat)
```

From our run, we get an MSE of 66.4 and R-squared of 65.84 with the auxiliary schools.

## Step 2: Generate Auxiliary Predictions for RCT Sample

It only takes one line of code to then generate predictions for the RCT sample: 

```{r, eval = F}
yhatrct <- predict(aux.rf, newdata = rct_dat)
```

We saved these predictions so that you don't have to re-run the auxiliary model, which takes a long time.

```{r, eval = F}
auxpreds <- data.frame(CAMPUS = rct_dat$CAMPUS, yhat = yhatrct)
save(auxpreds, file = "../temp/auxpred.Rdata")
```

## Step 3: RE-LOOP Estimate

### Adding Auxiliary Predictions

First, load the auxiliary predictions.

```{r}
load("../temp/auxpred.Rdata")
```

I will append the predictions to the RCT data to ensure that the predictions are in the correct order.

```{r}
rct_dat <- rct_dat %>%
  left_join(auxpreds, by = "CAMPUS")
```


```{r, echo = F}
rct_dat %>%
  select(CAMPUS, taks08, yhat) %>%
  slice(1:10)
```

We can get a sense of how well the auxiliary model transports to the RCT sample by looking at the R-squared of an OLS model using the predictions as a covariate:

```{r}
summary(lm(taks08 ~ yhat, data = rct_dat))$r.squared
```

This is a somewhat low R-squared, but it is still greater than the R-squared that you get with the pretest score alone:

```{r}
summary(lm(taks08~ premA, data = rct_dat))$r.squared
```

### Effect Estimate

The first option is to simply use the auxiliary predictions `yhat` as a covariate:

```{r}
loop(Y = rct_dat$taks08, Tr = rct_dat$TrBern, Z =rct_dat$yhat, pred = loop_ols)
```

Using the `pred = reloop` option, it will automatically determine whether to impute potential outcomes using `yhat`, or any RCT covariates that you give it in `Z`. 

So here, we specifiy `yhat = rct_dat$yhat` and separately `Z = rct_dat$premA`, the pretest score:

```{r}
loop(Y = rct_dat$taks08, Tr = rct_dat$TrBern, Z = rct_dat$premA, yhat = rct_dat$yhat, pred = reloop)
```

Given what we saw above, it makes sense that these point and standard error estimates are similar to just adjusting with the auxiliary predictions. 


And we can confirm that incorporating auxiliary information is more efficient than using the pretest score alone:

```{r}
loop(Y = rct_dat$taks08, Tr = rct_dat$TrBern, Z =rct_dat$premA, pred = loop_ols)
```


